The Effects of weather on maize yield: New evidence from Kenya

### 2019-01-15 writing/draft4/draft21.tex
yep, finilazing cross validaiton
writting correction

Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature including indicators of floods. These measures are listed in table~\ref{vars} in the appendix. The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation (CV) of seasonal precipitation, CV of average seasonal temperature (converted into Kelvin because the Celsius scale is interval and CV does not have any meaning for data measured on an interval scale),  maximum length of dry spell measured as number of consecutive dry days and the number of dry spells lasting for four days or more (for the purpose of this study we defined a dry day as a day when precipitation didn't exceed $1$mm). We did not include squared values of the average seasonal temperature in our preferred models as it was not significant and most of the past literature has not supported the evidence of the non-linear relationship between yield and average temperature.
\subsection{Statistical approach}\label{stats}

@@

To verify our results, we adapted an alternative step-down model building
approach using the Satterthwaite’s method to determine the p-values of the individual
t-tests \cite{lmerTest}. We started with a model which included the complete set of our weather measures in both fixed effects and random effects. In the first part of the procedure, the insignificant random effects were removed one by one. In each iteration, the insignificant variable with the highest p-value was removed. This was repeated until only the significant random effects remained. After this, we applied an analogous procedure to the fixed effects until we were left only with the significant variables. As recommended in literature, we considered the level of significance $\alpha=0.1$ for the random effects and $\alpha=0.05$~ for the fixed effects \cite{lmerTest}. This method led to the same model as the one obtained with our primary AIC-based method described above.

@@

>>>>

Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature including indicators of floods. These measures are listed in table~\ref{vars} in the appendix. The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation (CV) of seasonal precipitation, CV of average seasonal temperature (converted into Kelvin because the Celsius scale is interval and CV does not have any meaning for data measured on an interval scale),  maximum length of dry spell measured as number of consecutive dry days and the number of dry spells lasting for four days or more (for the purpose of this study we defined a dry day as a day when precipitation didn't exceed $1$ mm). We did not include squared values of the average seasonal temperature in our preferred model as it was not significant and most of the past literature has not supported the evidence of the non-linear relationship between yield and average temperature.
\subsection{Statistical approach}\label{stats}

@@

To verify our results, we adapted an alternative step-down model building
approach using the Satterthwaite’s method to determine the p-values of the individual
t-tests \cite{lmerTest}. We started with a model which included the complete set of our weather measures in both fixed effects and random effects. In the first part of the procedure, insignificant random effects were removed one by one. In each iteration, the insignificant variable with the highest p-value was removed. This was repeated until only significant random effects remained. After this, we applied an analogous procedure to the fixed effects until we were left only with significant variables. As recommended in literature, we considered the level of significance $\alpha=0.1$ for the random effects and $\alpha=0.05$~ for the fixed effects \cite{lmerTest}. This method led us to the same results as those obtained with our primary AIC-based method described above.

@@

\begin{center}
\begin{threeparttable}
\caption{\label{CVsc}Cross validation: Mean absolute prediction error scaled by IQR}
% files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
\begin{indented}
\item[]
\begin{tabular}{@{}lccc}
\\[-1em]
\br
\\[-1em]
\textbf{Sample:}&\textbf{All covariates}&\textbf{Only seasonal}&\textbf{Naive forecast}\\
\hline
{All counties}& $0.3811$&$0.3889$&$0.4277$\\
{ASAL}& $0.5300$&$0.5342$&$0.5835$\\
{non-ASAL}&  $0.2845$&$0.3858$&$0.4510$\\
\br
\end{tabular}
\begin{tablenotes}
\begin{footnotesize}
    \item \textit{Notes:} The coefficient of determination from generalised linear mixed-effects models. The statistic is based on \cite{Nakagawa2013}.  \\
        \begin{adjustwidth}{1cm}{}
    \item[a] The marginal $R^2$ is interpreted as the variance explained by the fixed effects
  \item[b] The conditional $R^2$ is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts)
     \end{adjustwidth}
     \end{footnotesize}
\end{tablenotes}

\begin{center}
\begin{threeparttable}
\caption{\label{CV}Cross validation: Mean absolute prediction error}
% files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
\begin{indented}
\item[]
\begin{tabular}{@{}lccc}
\\[-1em]
\br
\\[-1em]
\textbf{Sample:}&\textbf{All covariates}&\textbf{Only seasonal}&\textbf{Naive forecast}\\
\hline
{All counties}& $0.3599$&$0.3673$&$0.4040$\\
{ASAL}& $0.134$&$0.353$\\
{non-ASAL}&  $0.153$&$0.436$\\
\br
\end{tabular}
\begin{tablenotes}
\begin{footnotesize}
    \item \textit{Notes:} The coefficient of determination from generalised linear mixed-effects models. The statistic is based on \cite{Nakagawa2013}.  \\
        \begin{adjustwidth}{1cm}{}
    \item[a] The marginal $R^2$ is interpreted as the variance explained by the fixed effects
  \item[b] The conditional $R^2$ is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts)
     \end{adjustwidth}
     \end{footnotesize}
\end{tablenotes}
